Investigating the Safety of Different Toronto Neighbourhoods in 2011 Based on The Number of Crimes Reported

### 2025-06-18 paper/paper.qmd
updated paper to python

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(opendatatoronto)

cleaned_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
```

@@

This paper was created using statistical programming language R (@citeR) and was guided by (@rohan). Packages tidyverse (@citeTidyverse), knitr (@citeKnitr), ggplot2 (@citeGgplot2), and tibble (@citeTibble) were used as part of the code. The basis of the code was adapted from the book "Telling Stories with Data" (@tellingstorieswithdata).

@@

```{r}
#| label: fig-cleaned_data
#| fig-cap: Number of Different Types of Crimes By Neighbourhood 
#| echo: false

library(knitr)

ordered_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"), show_col_types = FALSE)

ordered_data |> 
  slice(1:10) |> 
  kable(
    col.names = c("Neighbourhood", "Assaults",
                  "Break & Enters", "Drug Arrests", "Hazardous Incidents", 
                  "Sexual Assaults", "Thefts", "Total Major Crime Incidents", 
                  "Vehicle Thefts"), digits = 1, booktabs = TRUE, linesep = ""
  )

@@

The data from the table (@fig-cleaned_data) show that the Waterfront Communities have the most total number of crime incidents in 2011. The remaining neighbourhoods in the top 10 include those on the left-most column labelled "Neighbourhood" in @fig-cleaned_data. The types of specific crimes that are viewed are in the column headings in row 1 of @fig-cleaned_data. These are the most common crimes that have occurred in the GTA in 2011, as represented by the data set.

@@

```{r}
#| label: fig-total_crimes_top_50
#| fig-cap: Top 50 Neighbourhoods in Total Major Crime Incidents
#| echo: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"), show_col_types = FALSE)

ggplot(data = analysis_data, mapping = aes(x = neighbourhood, y = total_major_crime_incidents)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "darkgrey") +
  theme_minimal() +
  labs(title = "Number of Total Major Crime Incidents By Neighbourhood
       (Top 50 Neighbourhoods)", 
       x = "Neighbourhood", 
       y = "Total Major Crime Incidents") +
  theme(axis.text.x = element_text(angle = 65, hjust = 1), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())

show_col_types = FALSE
```

@@

```{r}
#| label: fig-correlation
#| fig-cap: Relationship between Total Major Crime Incidents and Specific Crimes for The 50 Neighbourhoods with the Highest Number of Total Major Crime Incidents in Toronto in 2011
#| fig-subcap: ["Assaults", "Break and Enters", "Drug Arrests", "Hazardous Incidents", "Sexual Assaults", "Thefts", "Vehicle Thefts"]
#| layout-ncol: 3
#| echo: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"), show_col_types = FALSE)

ggplot(data = analysis_data, mapping = aes(x = total_major_crime_incidents, y = assaults)) +
  geom_point(color = "skyblue") +
  theme_minimal() +
  labs(title = "", 
       x = "Total Major Crime Incidents", 
       y = "Number of Assaults") +
  theme_minimal()

ggplot(data = analysis_data, mapping = aes(x = total_major_crime_incidents, y = break_and_enters)) +
  geom_point(color = "skyblue") +
  theme_minimal() +
  labs(title = "", 
       x = "Total Major Crime Incidents", 
       y = "Number of Break and Enters") +
  theme_minimal()

ggplot(data = analysis_data, mapping = aes(x = total_major_crime_incidents, y = drug_arrests)) +
  geom_point(color = "skyblue") +
  theme_minimal() +
  labs(title = "", 
       x = "Total Major Crime Incidents", 
       y = "Number of Drug Arrests") +
  theme_minimal()

ggplot(data = analysis_data, mapping = aes(x = total_major_crime_incidents, y = hazardous_incidents)) +
  geom_point(color = "skyblue") +
  theme_minimal() +
  labs(title = "", 
       x = "Total Major Crime Incidents", 
       y = "Number of Hazardous Incidents") +
  theme_minimal()

ggplot(data = analysis_data, mapping = aes(x = total_major_crime_incidents, y = sexual_assaults)) +
  geom_point(color = "skyblue") +
  theme_minimal() +
  labs(title = "", 
       x = "Total Major Crime Incidents", 
       y = "Number of Sexual Assault") +
  theme_minimal()

ggplot(data = analysis_data, mapping = aes(x = total_major_crime_incidents, y = thefts)) +
  geom_point(color = "skyblue") +
  theme_minimal() +
  labs(title = "", 
       x = "Total Major Crime Incidents", 
       y = "Number of Thefts") +
  theme_minimal()

ggplot(data = analysis_data, mapping = aes(x = total_major_crime_incidents, y = vehicle_thefts)) +
  geom_point(color = "skyblue") +
  theme_minimal() +
  labs(title = "", 
       x = "Total Major Crime Incidents", 
       y = "Number of Vehicle Thefts") +
  theme_minimal()
```

>>>>

```{python}
#| include: false
#| warning: false
#| message: false

import pandas as pd

cleaned_data = pd.read_csv("../data/analysis_data/analysis_data.csv")
```

@@

This paper was created using statistical programming language python (@python). Packages numpy (@numpy), matplotlib (@matplotlib) and pandas (@pandas) were used as part of the code.

@@

```{python}
#| label: fig-cleaned_data
#| fig-width: 10
#| fig-height: 2.5
#| echo: false

import pandas as pd
import matplotlib.pyplot as plt

# Load and prepare data
df = pd.read_csv("../data/analysis_data/analysis_data.csv").head(10)

# Shorten column names manually with \n
df = df.rename(columns={
    "neighbourhood": "Neighbourhood",
    "assaults": "Assaults",
    "break_and_enters": "Break\n& Enters",
    "drug_arrests": "Drug\nArrests",
    "hazardous_incidents": "Hazardous\nIncidents",
    "sexual_assaults": "Sexual\nAssaults",
    "thefts": "Thefts",
    "total_major_crime_incidents": "Total\nMajor\nCrimes",
    "vehicle_thefts": "Vehicle\nThefts"
})

# Convert to string (optional for display control)
table_data = df.values
column_labels = df.columns

# Create figure
fig, ax = plt.subplots(figsize=(12, 3))
ax.axis('off')

# Create the table
table = ax.table(
    cellText=table_data,
    colLabels=column_labels,
    loc='center',
    cellLoc='center'
)

table.auto_set_font_size(False)
table.set_fontsize(8)
table.scale(1, 1.5)  # Adjust height

plt.tight_layout()
plt.show()

@@

The data from the table (@fig-cleaned_data) show that the Waterfront Communities have the most total number of crime incidents in 2011. The remaining neighbourhoods in the top 10 include those on the left-most column labelled "Neighbourhood" in (@fig-cleaned_data). The types of specific crimes that are viewed are in the column headings in row 1 of (@fig-cleaned_data). These are the most common crimes that have occurred in the GTA in 2011, as represented by the data set.

@@

```{python}
#| label: fig-total_crimes_top_50
#| fig-cap: Top 50 Neighbourhoods in Total Major Crime Incidents
#| echo: false

import pandas as pd
import matplotlib.pyplot as plt

analysis_data = pd.read_csv("../data/analysis_data/analysis_data.csv")

plt.figure(figsize=(10, 6))
plt.bar(analysis_data["neighbourhood"], analysis_data["total_major_crime_incidents"],
        color="skyblue", edgecolor="darkgrey")
plt.title("Number of Total Major Crime Incidents By Neighbourhood\n(Top 50 Neighbourhoods)")
plt.xlabel("Neighbourhood")
plt.ylabel("Total Major Crime Incidents")
plt.xticks(rotation=65, ha='right')
plt.grid(False)
plt.tight_layout()
plt.show()

@@

```{python}
#| label: fig-correlation
#| fig-cap: Relationship between Total Major Crime Incidents and Specific Crimes for The 50 Neighbourhoods with the Highest Number of Total Major Crime Incidents in Toronto in 2011
#| fig-subcap: ["Assaults", "Break and Enters", "Drug Arrests", "Hazardous Incidents", "Sexual Assaults", "Thefts", "Vehicle Thefts"]
#| layout-ncol: 3
#| echo: false

import pandas as pd
import matplotlib.pyplot as plt

analysis_data = pd.read_csv("../data/analysis_data/analysis_data.csv")

fig, axes = plt.subplots(3, 3, figsize=(15, 12))
axes = axes.flatten()

# Plot variables (7 plots, 9 subplots to preserve layout)
plot_vars = [
    ("assaults", "Number of Assaults"),
    ("break_and_enters", "Number of Break and Enters"),
    ("drug_arrests", "Number of Drug Arrests"),
    ("hazardous_incidents", "Number of Hazardous Incidents"),
    ("sexual_assaults", "Number of Sexual Assault"),
    ("thefts", "Number of Thefts"),
    ("vehicle_thefts", "Number of Vehicle Thefts"),
]

for i, (col, y_label) in enumerate(plot_vars):
    ax = axes[i]
    ax.scatter(analysis_data["total_major_crime_incidents"], analysis_data[col], color="skyblue")
    ax.set_xlabel("Total Major Crime Incidents")
    ax.set_ylabel(y_label)
    ax.set_title("")  # Matches R version: no title
    ax.set_facecolor("white")

# Leave remaining subplots empty to match 3x3 layout
for i in range(len(plot_vars), 9):
    fig.delaxes(axes[i])

plt.tight_layout()
plt.show()
